---
title: "AdenylPred"
output:
  flexdashboard::flex_dashboard:
    theme: cerulean
    vertical_layout: fill
    logo: www/Lightning_Bolt_on_Circle.svg
    favicon: www/Lightning_Bolt_on_Circle.svg
    social: menu
runtime: shiny
---

```{r setup, include=FALSE}
library("xtable")
library("dplyr")
library("data.table")
library("shiny")
library("Biostrings")
library("tidyverse")
library("knitr")
library("kableExtra")
library("readxl")
library("DECIPHER")
library("ranger")
library("muscle")
library("DT")
```

```{r global}
## Read in the training set
## setwd("/Users/robi0916/Documents/Wageningen_UR/github/pallas/shiny_app")
source("src/extract_34_aa.r")
tab <- read_excel("data/combined_adenylate_forming_training_set_for_db_20191104.xlsx")
subkey <- read_excel("data/substrate_key.xlsx")
```

Home
=====================================
Column {.sidebar}
-----------------------------------------------------------------------
**Welcome to AdenylPred**
<br>
<br>
*Prediction tool for functional classification and substrate specificity of class I adenylate-forming enzymes*
<br>
<br>
AdenylPred uses supervised machine learning approaches to predict substrate specificity from amino acid sequences. 
<br>
<br>
This tool is the product of a collaboration between [Dr. Larry Wackett’s lab](https://cbs.umn.edu/wackett-lab/home) at the University of Minnesota and Dr. Marnix Medema's lab at Wageningen University and Research, the Netherlands. 
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
Enzymes in the class I adenylate-forming enzyme superfamily perform diverse functions including the adenylation of amino acids in nonribosomal peptide synthases (NRPS), luciferase-catalyzed light production, and the activation of short, medium, and long chain fatty acids with coenzyme A (CoA). 
<br>
<br>
Over the past decade, the adenylate-forming superfamily has expanded to include several new classes of enzymes including the fatty-acyl AMP ligases, aryl polyene adenylation domains, and β-lactone synthetases. Most ANL enzymes activate carboxylic acid groups with AMP then catalyze thioester bond formation with a sulfur acceptor such as a phosphopantetheine group.
<br>
<br>
As the number of sequences for adenylate-forming enzymes in databases far outnumber our capacity for experimental characterization, there is a need for adenylation enzyme substrate prediction tools. As a result, AdenylPred was developed using supervised machine learning approaches to predict substrate specificity from amino acid sequences. 
<br>
<br>
This tool is the product of a collaboration between [Dr. Larry Wackett’s lab](https://cbs.umn.edu/wackett-lab/home) at the University of Minnesota and Dr. Marnix Medema's lab at Wageningen University and Research, the Netherlands. 
<br>
<br>
We gratefully acknowledge support from a National Science Foundation graduate research fellowship to SLR funding this work.
<br>
<br>
Last updated: May 2, 2019

Column
-----------------------------------------------------------------------
```{r}
 # Input: Select a file ----
fileInput("file1", "Choose a FASTA File* \n" , multiple = FALSE)

# Horizontal line ----
tags$hr()

# Download button
downloadButton("downloadData", "Download", class = "download_this")
# tags$head(tags$style(".download_this{background-color:#add8e6;} .download_this{color: blue;} .download_this{height:40px;} .download_this{width:57px;}")) # background color and font color
#      
# Input: Select separator ----
radioButtons("seqtype", "Sequence type",
             choices = c(Nucleotide = "nuc",
                         Protein = "aa"),
             selected = "aa")

# Downloadable csv of selected dataset ----
downloadHandler(
    filename = function() {
      paste0("AdenylPred-", Sys.Date(), ".csv", sep = "")
    },
    content = function(file) {
      write.csv(mytab(), file, row.names = FALSE)
    }
  )

mytab <- reactive({
  
withProgress(message = 'Processing... Please wait', {

    # Input file
    req(input$file1)

    tryCatch(
      {
         if(input$seqtype == "aa") {
          sqs <- readAAStringSet(input$file1$datapath)
          if(length(sqs) > 100) {
            ret <- data.frame("Error: please input less than 100 sequences")
            colnames(ret) <- ""
            return(ret)
          }
         }
           else {
          sqs_dna <- readDNAStringSet(input$file1$datapath)
          if(!sqs_dna %in% c("A", "G", "T", "C", "N")) {
            ret <- data.frame("Error: not a valid DNA sequence")
            colnames(ret) <- ""
            return(ret)
          }
          sqs <- translate(sqs_dna, if.fuzzy.codon = "error")
          if(length(sqs) > 100) {
            ret <- data.frame("Error: please input less than 100 sequences")
            colnames(ret) <- ""
          }
        }
    

          query_fils <- sapply(1:length(sqs), function(x) {tempfile(pattern = "", fileext = ".fasta")})
          sapply(1:length(sqs), function(x) {writeXStringSet(sqs[x], query_fils[x])})
          extract_34_list <- lapply(1:length(sqs), function(x) { extract_34_aa(query_fils[x]) })
          extract_34_df <- data.frame(matrix(unlist(extract_34_list), nrow = length(sqs), byrow=T), 
                                      stringsAsFactors=FALSE)
          colnames(extract_34_df) <- as.character(fread("data/feature_names.txt", data.table = F)[,1])
          
          # Predict functional class
          rf_funct_class <- readRDS("data/rf_functional_class_1000trees_probability.rds")
          rf_funct_class_pred <- predict(rf_funct_class, data = extract_34_df, predict.all = F)
          rf_fc_pred <- rf_funct_class_pred$predictions
          res_fc_prob <- apply(rf_fc_pred, 1, max)
          res_fc <- tibble(colnames(rf_fc_pred)[apply(rf_fc_pred, 1, which.max)]) 
          res_long <- case_when(res_fc == "LACS" ~ "Long chain acyl-CoA synthetase",
                                res_fc == "MACS" ~ "Medium chain acyl-CoA synthetase",
                                res_fc == "SACS" ~ "Short chain acyl-CoA synthetase",
                                res_fc == "BLS" ~ "β-lactone synthetase",
                                res_fc == "LUCIFERASE" ~ "Firefly luciferase",
                                res_fc == "ARYL" ~ "Aryl-CoA ligase",
                                res_fc == "NRPS" ~ "Nonribosomal peptide synthetase",
                                res_fc == "FAAL" ~ "Fatty-acyl AMP ligase",
                                res_fc == "VLACSBILE" ~ "Very-long chain and bile acyl-CoA synthetase",
                                TRUE ~ "")

          # Predict small substrate
          rf_small_sub <- readRDS("data/rf_substrate_group_1000trees_probability.rds")
          rf_small_sub_pred <- predict(rf_small_sub, data = extract_34_df, predict.all = F)
          rf_ss_pred <- rf_small_sub_pred$predictions
          print(dim(rf_ss_pred))
          res_ss_prob <- apply(rf_ss_pred, 1, max)
          res_ss <- colnames(rf_ss_pred)[apply(rf_ss_pred, 1, which.max)]
          res_ss <- gsub("\\.", " ", res_ss)
          res_ss <- gsub("aa", " amino acid", res_ss)

          # Create a data frame of returns
          pred_df <- data.frame(names(sqs), res_long, round(res_fc_prob, 2), res_ss, round(res_ss_prob, 2), stringsAsFactors = F)
          colnames(pred_df) <- c("Query name", "Predicted functional class (FC)", "FC prediction probability",
                                 "Predicted substrate specificity (SS)", "SS prediction probability")
          return(pred_df)
},
  error = function(e) {
    # return a safeError if a parsing error occurs
    stop(safeError(e))
  }
  )
})
})

# DT::renderDataTable({ 
#   display_data <- DT::datatable(mytab())
#   # options = list(rowCallback = JS("function(row, data) {",
#   #                                 "var full_text = 'This row's values are:' + data[0] + ',' + data[1] + '...'", "$('td', row).attr('title', full_text);", "}")))
#   # options = list(rowCallback = JS("function(nRow, aData, iDisplayIndex, iDisplayIndexFull) {",
#   #                                 "var full_text = aData[1] + ',' + aData[2]",
#   #                                 "$('td:eq(1)', nRow).attr('title', full_text);",
#   #                                 "}")))
#                                   # columnDefs = list(className = 'dt-center', targets = "_all")), selection = 'single')
#                                   
#   })
renderTable({ mytab() }, align = 'c')

```
<br>

*Please note: This is a lightweight web app that cannot handle high computational loads. Please limit your analysis to 100 sequences. If you wish to analyze more sequences, please consider [downloading AdenylPred](https://github.com/serina-robinson/adenylpred) as a command line tool.


Substrate Groups
=====================================
Column {.sidebar data-width=350}
-----------------------------------------------------------------------
Due to lack of sufficient experimental training data for the exact substrate specificity of some adenylate-forming enzyme subfamilies, we are only able to predict at the level of "substrate group."

```{r}
my_list <- list()
my_list <- lapply(1:nrow(subkey), function(x) { my_list[[x]] <- x })
names(my_list) <- subkey$`Substrate group name`

radioButtons("radio_substrate_group", "Sequence type",
             choices = my_list)
```

Main
-----------------------------------------------------------------------

```{r}

rv <- reactiveValues(data = data.frame(subkey)[1,], img = "www/grp1.png")

observeEvent(input$radio_substrate_group, {
    ind <- as.numeric(input$radio_substrate_group)
    rv$data <- data.frame(subkey)[ind,]
    rv$img <- paste0("www/grp", ind, ".png")
})

output$img <- renderImage({
   list(src = rv$img)
}, deleteFile = FALSE)

renderTable({
  rv$data
})

imageOutput("img")
```


Resources
===================================== 
Column {.sidebar}
-----------------------------------------------------------------------

**Welcome to AdenylPred**
<br>
<br>
*Prediction tool for functional classification and substrate specificity of class I adenylate-forming enzymes*
<br>
<br>
Enzymes in the class I adenylate-forming enzyme superfamily perform diverse functions including the adenylation of amino acids in nonribosomal peptide synthases (NRPS), luciferase-catalyzed light production, and the activation of short, medium, and long chain fatty acids with coenzyme A (CoA). 
<br>
<br>
Over the past decade, the adenylate-forming superfamily has expanded to include several new classes of enzymes including the fatty-acyl AMP ligases, aryl polyene adenylation domains, and β-lactone synthetases. Most ANL enzymes activate carboxylic acid groups with AMP then catalyze thioester bond formation with a sulfur acceptor such as a phosphopantetheine group.
<br>
<br>
As the number of sequences for adenylate-forming enzymes in databases far outnumber our capacity for experimental characterization, there is a need for adenylation enzyme substrate prediction tools. As a result, AdenylPred was developed using supervised machine learning approaches to predict substrate specificity from amino acid sequences. 
<br>
<br>
This tool is the product of a collaboration between [Dr. Larry Wackett’s lab](https://cbs.umn.edu/wackett-lab/home) at the University of Minnesota and Dr. Marnix Medema's lab at Wageningen University and Research, the Netherlands. 
<br>
<br>
We gratefully acknowledge support from a National Science Foundation graduate research fellowship to SLR funding this work.
<br>
<br>
Last updated: April 29, 2019

Inputs
-----------------------------------------------------------------------
**Relevant Scientific Publications on Adenlyate-forming enzymes**
  
1. Gulick, A. M. (2009). [Conformational Dynamics in the Acyl-CoA Synthetases, Adenylation Domains of Non-Ribosomal Peptide Synthetases, and Firefly Luciferase.](https://pubs.acs.org/doi/10.1021/cb900156h) *ACS Chemical Biology* 2009, 4(10), 811-827. [doi: 10.1021/cb900156h.](https://pubs.acs.org/doi/10.1021/cb900156h)
  
**Publications Specific to β-Lactone Synthetases**

1. Robinson, SL, Christenson, JK, Richman, JE, Jenkins, DJ, Neres, J, Fonseca, DR, Aldrich, CC, Wackett, LP (2019). [Mechanism of a Standalone β-Lactone Synthetase: New Continuous Assay for a Widespread ANL Superfamily Enzyme.](https://onlinelibrary.wiley.com/doi/abs/10.1002/cbic.201800821) *ChemBioChem* 2019, [Epub ahead of print]. [doi: 10.1002/cbic.201800821.](https://onlinelibrary.wiley.com/doi/abs/10.1002/cbic.201800821)    

2. Christenson JK, Richman JE, Jensen MR, Neufeld JY, Wilmot CM, Wackett LP. [β-Lactone Synthetase Found in the Olefin Biosynthesis Pathway.](http://pubs.acs.org/doi/abs/10.1021/acs.biochem.6b01199) *Biochemistry* 2017, 56(2):348-51.  [doi:10.1021/acs.biochem.6b01199.](http://pubs.acs.org/doi/abs/10.1021/acs.biochem.6b01199)    

Database
===================================== 

Inputs {.sidebar}
-----------------------------------------------------------------------
<br>
<br>
Training set database of adenylate-forming enzymes paired with substrate. 
```{r}

      selectizeInput("small_substrate_group", "Search by substrate group:",
                    choices = tab$small_substrate_group,
                    selected = "luciferase", multiple = TRUE
      )
      
      selectizeInput("functional_class", "Search by enzyme functional class:",
                    choices = tab$functional_class,
                    selected = NULL, multiple = TRUE
      )

       selectizeInput("likely_substrate", "Search by specific substrate:",
                    choices = tab$substrate,
                    selected = NULL, multiple = TRUE
      )
      
      selectizeInput("organism", "Search by organism:",
                   choices = tab$organism,
                   selected = NULL, multiple = TRUE
      )
```

Outputs
-----------------------------------------------------------------------
```{r}   
 DT::renderDataTable({
      srch <- tab[tab$substrate %in% input$likely_substrate,]
      funct <- tab[tab$functional_class %in% input$functional_class,]
      cmpd <- tab[tab$small_substrate_group %in% input$small_substrate_group,]
      tax <- tab[tab$organism %in% input$organism,]
      out <- data.frame(rbind(cmpd, tax, funct, srch))
      dedup <- out[!duplicated(out),]

      return(xtable(dedup))
    })
```   

About
=====================================
Main
-----------------------------------------------------------------------
**Welcome to AdenylPred**
<br>
<br>
This tool is the product of a collaboration between [Dr. Larry Wackett’s lab](https://cbs.umn.edu/wackett-lab/home) at the University of Minnesota and Dr. Marnix Medema's lab at Wageningen University and Research, the Netherlands.
<br>
<br>
This material is based upon work supported by the National Science Foundation Graduate Research Fellowship under Grant No. 00039202 to SLR.
<br>
<br> 
Disclaimer: Any opinions, findings, and conclusions or recommendations expressed in this material are those of the author(s) and do not necessarily reflect the views of the National Science Foundation.
<br>
<br>
**Questions, Comments, or Suggestions?**
Please [contact us](mailto:www.robinsonserinalee@gmail.com)!


